---
output: 
  github_document:
  pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# discrtr

<!-- badges: start -->
<!-- badges: end -->

Package {discrtr} is a companion to the book [Discrete Choice Analysis with R](https://link.springer.com/book/10.1007/978-3-031-20719-8). It includes data sets and other supporting materials used in the book.

## Installation

You can install the development version of discrtr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("paezha/discrtr")
```

## Example

One data set included in the package relates to commuting in a university setting. Analysis of this data was reported by Whalen, P\'{a}ez, and Carrasco [(2012)](https://doi.org/10.1016/j.jtrangeo.2013.06.008). The wide version of the data table can be loaded as follows:

```{r load-data}
library(discrtr)
data("mc_commute_wide")
```

This is a quick summary of the travel time variables in the table:

```{r summary, message=FALSE}
library(dplyr)
mc_commute_wide %>% 
  dplyr::select(starts_with("time.")) %>%
  summary()
```

There are four modes of transportation in this data set, but not all modes were available to every respondent, hence the NA's. A missing value for any of the time variables indicates that the mode was not part of the choice set for the corresponding traveller.
