---
output: 
  github_document:
  pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# discrtr

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/discrtr)](https://CRAN.R-project.org/package=discrtr)

[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)

[![GitHub Stars](https://img.shields.io/github/stars/paezha/discrtr?style=social)](https://github.com/paezha/discrtr/stargazers)

![GitHub search hit counter: R](https://img.shields.io/github/search/paezha/discrtr/R)  
![GitHub search hit counter: data](https://img.shields.io/github/search/paezha/discrtr/data)

![GitHub issues](https://img.shields.io/github/issues/paezha/discrtr)
![GitHub release](https://img.shields.io/github/release-date/paezha/discrtr)
![GitHub commit activity](https://img.shields.io/github/commit-activity/y/paezha/discrtr)
![GitHub last commit](https://img.shields.io/github/last-commit/paezha/discrtr)
[![](https://cranlogs.r-pkg.org/badges/discrtr)](https://cran.r-project.org/package=discrtr)
<!-- badges: end -->

Package {discrtr} is a companion to the book [Discrete Choice Analysis with R](https://link.springer.com/book/10.1007/978-3-031-20719-8). It includes templates with code and data sets used in the book.

## Installation

Install the released version of usethis from CRAN:
```r
install.packages("discrtr")
```

Or you can install the development version of discrtr from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("paezha/discrtr")
```

## Example

One data set included in the package relates to commuting in a university setting. Analysis of this data was reported by Whalen, Páez, and Carrasco [(2012)](https://doi.org/10.1016/j.jtrangeo.2013.06.008). The wide version of the data table can be loaded as follows:

```{r load-data}
library(discrtr)
data("mc_commute_wide")
```

This is a quick summary of the travel time variables in the table:

```{r summary, message=FALSE}
library(dplyr)
mc_commute_wide %>% 
  dplyr::select(starts_with("time.")) %>%
  summary()
```

There are four modes of transportation in this data set, but not all modes were available to every respondent, hence the NA's. A missing value for any of the time variables indicates that the mode was not part of the choice set for the corresponding traveller.
