---
title: "R Notebook"
---

## Introduction

In this notebook we prepare a data object with the table used in Lavery, T. A., Antonio Páez, and Pavlos S. Kanaroglou. "Driving out of choices: An investigation of transport modality in a university sample." Transportation research part A: policy and practice 57 (2013): 37-46.

## Preliminaries

Clear environment:
```{r}
rm(list = ls())
```

Load packages used in the notebook:
```{r message = FALSE}
library(readxl)
library(tidyverse)
#library(MASS)
#library(reshape2)
#library(plyr)
#library(kableExtra)
#library(gridExtra)
```

Load dataframe:
```{r}
mc_modality <- read_excel("input-data-files/Database+Geocodes.xlsx")
### Hamilton Ward boundaries
```

Convert to factors:
```{r}
mc_modality <- mc_modality %>% 
  select(-c(Longitude, Latitude, XUTM, YUTM)) %>%
  rename(id = ID,
         choice = MODE,
         gender = Gender,
         Community = COMMUNITY,
         ActiveNeigh = ACTIVE_NEIGHBORHOOD,
         Neighbors = NEIGHBORS,
         Safe_Walk= SAFE_WALK,
         Shops_Important = SHOPS_IMPORTANT,
         TravelAlone = TRAVEL_ALONE) %>%
  mutate(choice = case_when(choice == ""),
         gender = case_when(gender == "FEMALE" ~ "Woman",
                                                         gender == "MALE" ~ "Man"),
                                      COMMUNITY = factor(COMMUNITY, 
                                       levels = c("STRONGLY DISAGREE", "DISAGREE", "NEUTRAL", "AGREE", "STRONGLY AGREE"), ordered = TRUE),
                    ActiveNeigh = factor(ActiveNeigh, 
                                       levels = c("STRONGLY DISAGREE", "DISAGREE", "NEUTRAL", "AGREE", "STRONGLY AGREE"), ordered = TRUE),
                    NEIGHBORS = factor(NEIGHBORS, 
                                       levels = c("STRONGLY DISAGREE", "DISAGREE", "NEUTRAL", "AGREE", "STRONGLY AGREE"), ordered = TRUE),
                    SAFE_WALK = factor(SAFE_WALK, 
                                       levels = c("STRONGLY DISAGREE", "DISAGREE", "NEUTRAL", "AGREE", "STRONGLY AGREE"), ordered = TRUE),
                    SHOPS_IMPORTANT = factor(SHOPS_IMPORTANT, 
                                       levels = c("STRONGLY DISAGREE", "DISAGREE", "NEUTRAL", "AGREE", "STRONGLY AGREE"), ordered = TRUE),
                    TRAVEL_ALONE = factor(TRAVEL_ALONE, 
                                       levels = c("STRONGLY DISAGREE", "DISAGREE", "NEUTRAL", "AGREE", "STRONGLY AGREE"), ordered = TRUE),
                    MODE = factor(MODE, 
                                       levels = c("WALK", "HSR", "CAR")),
                    Gender = factor(Gender, 
                                       levels = c("FEMALE", "MALE")),
                    License = factor(License, 
                                       levels = c("YES", "NO")),
                    Vehicle = factor(Vehicle, 
                                       levels = c("YES", "NO")),
                    Visa = factor(Visa, 
                                       levels = c("CANADIAN", "INTERNATIONAL"))
)
```

Scale income variables:
```{r}
mc_modality <- mc_modality %>% mutate(MedIncome = MedIncome/10000, MHI = MHI/10000)
```

Scale density variables:
```{r}
mc_modality <- mc_modality %>% mutate(EMPLOY_DEN = EMPLOY_DEN/10000, POP_DENSIT = POP_DENSIT/10000)
```

Save data:
```{r}
usethis::use_data(mc_modality)
```

